Empowering the use of preprints in ecology and evolution: an exploration of the first 1000 *EcoEvoRxiv* preprints

### 2023-10-12 ms/ms.qmd
Finished first draft of intro

Preprints attempt to break down barriers to scientific publishing by: 1) increasing the speed, visibility and output of research findings; 2) encouraging constructive and critical feedback from a research community; 3) helping establish precedence of research findings; and 4) removing financial barriers to publication [@Proulx2013 and others]. Importantly, *EcoEvoRxiv* is a community driven preprint server that has paved the way for new initiatives, including accepting multilingual preprints (i.e., Spanish and Portuguese), recognising and accepting registered reports and promoting the use of peer review and community discussion to improve the quality of preprints and speed up their peer-reviewed publication (e.g., Peer Community In -- PCI).

Preprint servers, such as *EcoEvoRxiv*, can empower researchers to make their findings more accessible, open and transparent but only if they are used well. Significant barriers to preprint use still exist globally – from worry about XXX to XXX. Nonetheless, we lack a systematic understanding of the factors that influence preprint use in ecology and evolution which can help inform future initiatives and allow us to begin to break down barriers to their use in our field.

>>>>

Preprints attempt to break down barriers to scientific publishing by: 1) increasing the speed, visibility and output of research findings; 2) encouraging constructive and critical feedback from a research community; 3) helping establish precedence of research findings; and 4) removing financial barriers to publication [e.g., @Proulx2013]. Importantly, *EcoEvoRxiv* is a community driven preprint server that has paved the way for new initiatives, including accepting multilingual preprints, recognising and accepting registered reports and non-traditional research reports, and promoting peer review and community discussion to improve the quality of preprints and speed up their peer-reviewed publication (e.g., Peer Community In -- PCI).

Preprint servers, such as *EcoEvoRxiv*, can empower researchers to make their findings more accessible, open and transparent but only if they are used well. Significant barriers to preprint use still exist globally – from lack of clarity around preprint policies at journals [@Klebel2020] to a stigma that preprints are viewed by a research community to be of poor quality [@Chiarelli2019]. Nonetheless, we lack a systematic understanding of the factors that influence preprint use in ecology and evolution which can help inform future initiatives and allow us to begin to break down barriers to their use in our field.
